---
title: "Analysis of tRNA Pool"
author: "Vatsal"
date: "12/8/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
```
things to do with data...

make super dataset
compare average error rates between two species - check differences
t test(google)=
anova(compare mean across 2+ groups) -> compare mean of all organism. compare depending on psycro, mneso, and thermo

look elongation rates(Rc) and missense(eM)


add more organism -> diversity
Add Pseudoalteromonas haloplanktis TAC125
Add organisms with large tRNA gene #
```{r scatter plot}

```
sum up middle colum(tRNA)
and compare them
do pyscrophile have more tRNA gene compared to other enviorment


```{r ecoli vs shewSedi(psycrophile)}


our <- read_tsv("Rates/Psychrophile/flavPsyc_3.tsv")
key <- read_tsv("Rates/Mesophile/ecoli_tRNA_2010.tsv")

combine <- our %>% left_join(key,by=c("AA","Codon"),suffix = c("_flavPsyc", "_ecoli"))
p <- ggplot(combine, aes(x=Rc_flavPsyc,y=Rc_ecoli)) + geom_point() + geom_abline(intercept = 0,slope = 1) + xlab("Elongation Rates \n flavPsyc") + ylab("Elongation Rates \n ecoli")

c <- our %>% left_join(key,by=c("AA","Codon"),suffix = c("_flavPsyc", "_ecoli"))
pl <- ggplot(c, aes(x=eM_flavPsyc,y=eM_ecoli)) + geom_point() + geom_abline(intercept = 0,slope = 1) + xlab("Missense Rates \n flavPsyc") + ylab("Missense Rates \n ecoli")

p
pl


```


```{r ecoli vs shewSedi(psycrophile)}


our <- read_tsv("Rates/Psychrophile/shewSedi_HAW_EB3.tsv")
key <- read_tsv("Rates/Mesophile/ecoli_tRNA_2010.tsv")

combine <- our %>% left_join(key,by=c("AA","Codon"),suffix = c("_shewSedi", "_ecoli"))
p <- ggplot(combine, aes(x=Rc_shewSedi,y=Rc_ecoli)) + geom_point() + geom_abline(intercept = 0,slope = 1) + xlab("Elongation Rates \n shewSedi") + ylab("Elongation Rates \n ecoli")

c <- our %>% left_join(key,by=c("AA","Codon"),suffix = c("_shewSedi", "_ecoli"))
pl <- ggplot(c, aes(x=eM_shewSedi,y=eM_ecoli)) + geom_point() + geom_abline(intercept = 0,slope = 1) + xlab("Missense Rates \n shewSedi") + ylab("Missense Rates \n ecoli")

p
pl


```



```{r ecoli vs shewLoih(psycrophile)}


our <- read_tsv("Rates/Psychrophile/shewLoihPV4.tsv")
key <- read_tsv("Rates/Mesophile/ecoli_tRNA_2010.tsv")

combine <- our %>% left_join(key,by=c("AA","Codon"),suffix = c("_shewLoih", "_ecoli"))
p <- ggplot(combine, aes(x=Rc_shewLoih,y=Rc_ecoli)) + geom_point() + geom_abline(intercept = 0,slope = 1) + xlab("Elongation Rates \n shewLoih") + ylab("Elongation Rates \n ecoli")

c <- our %>% left_join(key,by=c("AA","Codon"),suffix = c("_shewLoih", "_ecoli"))
pl <- ggplot(c, aes(x=eM_shewLoih,y=eM_ecoli)) + geom_point() + geom_abline(intercept = 0,slope = 1) + xlab("Missense Rates \n shewLoih") + ylab("Missense Rates \n ecoli")

p
pl


```

```{r ecoli vs desuPsyc(psycrophile)}


our <- read_tsv("Rates/Psychrophile/desuPsyc_LSV54.tsv")
key <- read_tsv("Rates/Mesophile/ecoli_tRNA_2010.tsv")

combine <- our %>% left_join(key,by=c("AA","Codon"),suffix = c("_desuPsyc", "_ecoli"))
p <- ggplot(combine, aes(x=Rc_desuPsyc,y=Rc_ecoli)) + geom_point() + geom_abline(intercept = 0,slope = 1) + xlab("Elongation Rates \n desuPsyc") + ylab("Elongation Rates \n ecoli")

c <- our %>% left_join(key,by=c("AA","Codon"),suffix = c("_desuPsyc", "_ecoli"))
pl <- ggplot(c, aes(x=eM_desuPsyc,y=eM_ecoli)) + geom_point() + geom_abline(intercept = 0,slope = 1) + xlab("Missense Rates \n desuPsyc") + ylab("Missense Rates \n ecoli")

p
pl


```





















```{r ecoli vs geoTher(Thermophile)}
our <- read_tsv("Rates/Thermophile/geobTher_NG80_2.tsv")
key <- read_tsv("Rates/Mesophile/ecoli_tRNA_2010.tsv")

combine <- our %>% left_join(key,by=c("AA","Codon"),suffix = c("_geobTher", "_ecoli"))
p <- ggplot(combine, aes(x=Rc_geobTher,y=Rc_ecoli)) + geom_point() + geom_abline(intercept = 0,slope = 1) + xlab("Elongation Rates \n geother") + ylab("Elongation Rates \n ecoli")

c <- our %>% left_join(key,by=c("AA","Codon"),suffix = c("_geobTher", "_ecoli"))
pl <- ggplot(c, aes(x=eM_geobTher,y=eM_ecoli)) + geom_point() + geom_abline(intercept = 0,slope = 1) + xlab("Missense Rates \n geobTher") + ylab("Missense Rates \n ecoli")


p
pl
```

```{r ecoli vs heliMode(Thermophile)}
our <- read_tsv("Rates/Thermophile/heliMode_ICE1_ATCC51547.tsv")
key <- read_tsv("Rates/Mesophile/ecoli_tRNA_2010.tsv")

combine <- our %>% left_join(key,by=c("AA","Codon"),suffix = c("_heliMode", "_ecoli"))
p <- ggplot(combine, aes(x=Rc_heliMode,y=Rc_ecoli)) + geom_point() + geom_abline(intercept = 0,slope = 1) + xlab("Elongation Rates \n heliMode") + ylab("Elongation Rates \n ecoli")

c <- our %>% left_join(key,by=c("AA","Codon"),suffix = c("_heliMode", "_ecoli"))
pl <- ggplot(c, aes(x=eM_heliMode,y=eM_ecoli)) + geom_point() + geom_abline(intercept = 0,slope = 1) + xlab("Missense Rates \n heliMode") + ylab("Missense Rates \n ecoli")


p
pl
```



```{r ecoli vs hydrSp(Thermophile)}
our <- read_tsv("Rates/Thermophile/hydrSp_Y04AAS1.tsv")
key <- read_tsv("Rates/Mesophile/ecoli_tRNA_2010.tsv")

combine <- our %>% left_join(key,by=c("AA","Codon"),suffix = c("_hydrSp", "_ecoli"))
p <- ggplot(combine, aes(x=Rc_hydrSp,y=Rc_ecoli)) + geom_point() + geom_abline(intercept = 0,slope = 1) + xlab("Elongation Rates \n hydrSp") + ylab("Elongation Rates \n ecoli")

c <- our %>% left_join(key,by=c("AA","Codon"),suffix = c("_hydrSp", "_ecoli"))
pl <- ggplot(c, aes(x=eM_hydrSp,y=eM_ecoli)) + geom_point() + geom_abline(intercept = 0,slope = 1) + xlab("Missense Rates \n hydrSp") + ylab("Missense Rates \n ecoli")


p
pl
```

```{r ecoli vs therSp(Thermophile)}
our <- read_tsv("Rates/Thermophile/therSp_X514.tsv")
key <- read_tsv("Rates/Mesophile/ecoli_tRNA_2010.tsv")

combine <- our %>% left_join(key,by=c("AA","Codon"),suffix = c("_therSp", "_ecoli"))
p <- ggplot(combine, aes(x=Rc_therSp,y=Rc_ecoli)) + geom_point() + geom_abline(intercept = 0,slope = 1) + xlab("Elongation Rates \n therSp") + ylab("Elongation Rates \n ecoli")

c <- our %>% left_join(key,by=c("AA","Codon"),suffix = c("_therSp", "_ecoli"))
pl <- ggplot(c, aes(x=eM_therSp,y=eM_ecoli)) + geom_point() + geom_abline(intercept = 0,slope = 1) + xlab("Missense Rates \n therSp") + ylab("Missense Rates \n ecoli")


p
pl
```























```{r baciSubt2 vs flavPysc(psycrophile)}


our <- read_tsv("Rates/Psychrophile/flavPsyc_3.tsv")
key <- read_tsv("Rates/Mesophile/baciSubt2_tRNA.tsv")

combine <- our %>% left_join(key,by=c("AA","Codon"),suffix = c("_flavPsyc", "_baciSubt2"))
p <- ggplot(combine, aes(x=Rc_flavPsyc,y=Rc_baciSubt2)) + geom_point() + geom_abline(intercept = 0,slope = 1) + xlab("Elongation Rates \n flavPsyc") + ylab("Elongation Rates \n baciSubt2")

c <- our %>% left_join(key,by=c("AA","Codon"),suffix = c("_flavPsyc", "_baciSubt2"))
pl <- ggplot(c, aes(x=eM_flavPsyc,y=eM_baciSubt2)) + geom_point() + geom_abline(intercept = 0,slope = 1) + xlab("Missense Rates \n flavPsyc") + ylab("Missense Rates \n baciSubt2")

p
pl


```


```{r baciSubt2 vs shewSedi(psycrophile)}


our <- read_tsv("Rates/Psychrophile/shewSedi_HAW_EB3.tsv")
key <- read_tsv("Rates/Mesophile/baciSubt2_tRNA.tsv")

combine <- our %>% left_join(key,by=c("AA","Codon"),suffix = c("_shewSedi", "_baciSubt2"))
p <- ggplot(combine, aes(x=Rc_shewSedi,y=Rc_baciSubt2)) + geom_point() + geom_abline(intercept = 0,slope = 1) + xlab("Elongation Rates \n shewSedi") + ylab("Elongation Rates \n baciSubt2")

c <- our %>% left_join(key,by=c("AA","Codon"),suffix = c("_shewSedi", "_baciSubt2"))
pl <- ggplot(c, aes(x=eM_shewSedi,y=eM_baciSubt2)) + geom_point() + geom_abline(intercept = 0,slope = 1) + xlab("Missense Rates \n shewSedi") + ylab("Missense Rates \n baciSubt2")

p
pl


```



```{r baciSubt2 vs shewLoih(psycrophile)}


our <- read_tsv("Rates/Psychrophile/shewLoihPV4.tsv")
key <- read_tsv("Rates/Mesophile/baciSubt2_tRNA.tsv")

combine <- our %>% left_join(key,by=c("AA","Codon"),suffix = c("_shewLoih", "_baciSubt2"))
p <- ggplot(combine, aes(x=Rc_shewLoih,y=Rc_baciSubt2)) + geom_point() + geom_abline(intercept = 0,slope = 1) + xlab("Elongation Rates \n shewLoih") + ylab("Elongation Rates \n baciSubt2")

c <- our %>% left_join(key,by=c("AA","Codon"),suffix = c("_shewLoih", "_baciSubt2"))
pl <- ggplot(c, aes(x=eM_shewLoih,y=eM_baciSubt2)) + geom_point() + geom_abline(intercept = 0,slope = 1) + xlab("Missense Rates \n shewLoih") + ylab("Missense Rates \n baciSubt2")

p
pl


```

```{r baciSubt2 vs desuPsyc(psycrophile)}


our <- read_tsv("Rates/Psychrophile/desuPsyc_LSV54.tsv")
key <- read_tsv("Rates/Mesophile/ecoli_tRNA_2010.tsv")

combine <- our %>% left_join(key,by=c("AA","Codon"),suffix = c("_desuPsyc", "_baciSubt2"))
p <- ggplot(combine, aes(x=Rc_desuPsyc,y=Rc_baciSubt2)) + geom_point() + geom_abline(intercept = 0,slope = 1) + xlab("Elongation Rates \n desuPsyc") + ylab("Elongation Rates \n baciSubt2")

c <- our %>% left_join(key,by=c("AA","Codon"),suffix = c("_desuPsyc", "_baciSubt2"))
pl <- ggplot(c, aes(x=eM_desuPsyc,y=eM_baciSubt2)) + geom_point() + geom_abline(intercept = 0,slope = 1) + xlab("Missense Rates \n desuPsyc") + ylab("Missense Rates \n baciSubt2")

p
pl


```





















```{r baciSubt2 vs geoTher(Thermophile)}
our <- read_tsv("Rates/Thermophile/geobTher_NG80_2.tsv")
key <- read_tsv("Rates/Mesophile/baciSubt2_tRNA.tsv")

combine <- our %>% left_join(key,by=c("AA","Codon"),suffix = c("_geobTher", "_baciSubt2"))
p <- ggplot(combine, aes(x=Rc_geobTher,y=Rc_baciSubt2)) + geom_point() + geom_abline(intercept = 0,slope = 1) + xlab("Elongation Rates \n geother") + ylab("Elongation Rates \n baciSubt2")

c <- our %>% left_join(key,by=c("AA","Codon"),suffix = c("_geobTher", "_baciSubt2"))
pl <- ggplot(c, aes(x=eM_geobTher,y=eM_baciSubt2)) + geom_point() + geom_abline(intercept = 0,slope = 1) + xlab("Missense Rates \n geobTher") + ylab("Missense Rates \n baciSubt2")


p
pl
```

```{r baciSubt2 vs geoTher(Thermophile)}
our <- read_tsv("Rates/Thermophile/heliMode_ICE1_ATCC51547.tsv")
key <- read_tsv("Rates/Mesophile/baciSubt2_tRNA.tsv")

combine <- our %>% left_join(key,by=c("AA","Codon"),suffix = c("_heliMode", "_baciSubt2"))
p <- ggplot(combine, aes(x=Rc_heliMode,y=Rc_baciSubt2)) + geom_point() + geom_abline(intercept = 0,slope = 1) + xlab("Elongation Rates \n heliMode") + ylab("Elongation Rates \n baciSubt2")

c <- our %>% left_join(key,by=c("AA","Codon"),suffix = c("_heliMode", "_baciSubt2"))
pl <- ggplot(c, aes(x=eM_heliMode,y=eM_baciSubt2)) + geom_point() + geom_abline(intercept = 0,slope = 1) + xlab("Missense Rates \n heliMode") + ylab("Missense Rates \n baciSubt2")


p
pl
```



```{r baciSubt2 vs hydrSp(Thermophile)}
our <- read_tsv("Rates/Thermophile/hydrSp_Y04AAS1.tsv")
key <- read_tsv("Rates/Mesophile/baciSubt2_tRNA.tsv")

combine <- our %>% left_join(key,by=c("AA","Codon"),suffix = c("_hydrSp", "_baciSubt2"))
p <- ggplot(combine, aes(x=Rc_hydrSp,y=Rc_baciSubt2)) + geom_point() + geom_abline(intercept = 0,slope = 1) + xlab("Elongation Rates \n hydrSp") + ylab("Elongation Rates \n baciSubt2")

c <- our %>% left_join(key,by=c("AA","Codon"),suffix = c("_hydrSp", "_baciSubt2"))
pl <- ggplot(c, aes(x=eM_hydrSp,y=eM_baciSubt2)) + geom_point() + geom_abline(intercept = 0,slope = 1) + xlab("Missense Rates \n hydrSp") + ylab("Missense Rates \n baciSubt2")


p
pl
```

```{r baciSubt2 vs therSp(Thermophile)}
our <- read_tsv("Rates/Thermophile/therSp_X514.tsv")
key <- read_tsv("Rates/Mesophile/baciSubt2_tRNA.tsv")

combine <- our %>% left_join(key,by=c("AA","Codon"),suffix = c("_therSp", "_baciSubt2"))
p <- ggplot(combine, aes(x=Rc_therSp,y=Rc_baciSubt2)) + geom_point() + geom_abline(intercept = 0,slope = 1) + xlab("Elongation Rates \n therSp") + ylab("Elongation Rates \n baciSubt2")

c <- our %>% left_join(key,by=c("AA","Codon"),suffix = c("_therSp", "_baciSubt2"))
pl <- ggplot(c, aes(x=eM_therSp,y=eM_baciSubt2)) + geom_point() + geom_abline(intercept = 0,slope = 1) + xlab("Missense Rates \n therSp") + ylab("Missense Rates \n baciSubt2")


p
pl
```






